\Huge Conferences in the Era of Expensive Carbon

### 2019-06-04 main.tex
Checkpoint work on writing

\title{\Huge Conferences and Cost of Carbon}
\author{Benjamin C. Pierce (University of Pennsylvania) \\
Michael Hicks (University of Maryland) \\
Crista Lopes (University California, Irvine) \\
Jens Palsberg (University of California, Los Angeles)}

@@

A broad scientific consensus warns that the earth is warming, due to human
emissions of carbon dioxide and other greenhouse gases.
% (Figure \ref{fig:warming}).
This is not a problem we can leave to future
generations: we are up against it now. The UN's Intergovernmental
Panel on Climate Change (IPCC) recently recommended a 40\% decrease in
emissions by 2030 to avoid irreversible damage~\cite{IPCC-report}.
% (Figure~\ref{fig:IPCC-cover}).
Achieving reductions on this scale will
require urgent and sustained commitment at all levels of society, including
non-governmental organizations such as universities, companies, and scientific
societies like ACM.

Indeed, ACM arguably has an especially important role to play, since one of
its main activities is organizing international
conferences, which produce significant emissions of greenhouse gases,
particularly from air travel.   A single round-trip flight from
Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
dioxide (\COtwoE) per passenger~\cite{carboncalculator}\iflater\bcp{Trim
  citation?}\fi---a substantial fraction of the {\em total} yearly emissions
for an average resident of the US (16.5 tons) or Europe (7
tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put in some
  figures about the footprints of a few other things that people might think
  of reducing---driving, eating meat, upgrading their fridge...)}\fi{}
Moreover, these emissions have no near-term 
technological fix because jet fuel is difficult to replace with
renewable energy sources~\cite{elec-air}. 
\iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
of conferences will require a reduction in the footprint of the attendees'
air travel.}\fi

On the other hand, conferences obviously also create value.  Science is a
fundamentally social process, and the conference system accelerates research
through opportunities for high-bandwidth interaction, direct dissemination
of results, network building, and serendipitous cross-fertilization.

This leads to a puzzle: How do we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  This seems an extremely difficult question to
answer: the harm and the compensating benefits cannot be directly
compared---they are apples and oranges.  
%
Fortunately, there is a well-established mechanism for comparing apples and
oranges---indeed, for comparing almost anything: the magic of markets.  By
mapping choices to currency amounts, markets facilitate individual valuation
of costs and benefits across vastly different activities.\bcp{Give a simple,
  concrete example.} For this reason, many policy experts advocate some form
of carbon pricing, which places a monetary cost on greenhouse gas emissions,
giving a market signal that changes aggregate behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}
Exactly how this should be done is a challenging question, but it has a
clear success 
criterion: The cost should be set so that the world arrives at zero
Emissions within the timeframe that climate scientists tell us is safe.

In the long term, this cost must be imposed from outside, by governments.
Indeed, many countries and states (e.g., California) have already adopted
carbon pricing policies; more will follow.
%
Moreover, although the current market price of carbon is still quite low
(around \$15 / ton), experts agree that this will need to rise significantly
over time to achieve the required effect.  Bottom line: carbon is going to
be expensive.

Expecting that this is coming, what should we rationally do now to ensure
that ACM, and computer-science research broadly, continue to thrive in a
world where air travel is more expensive?  Believing that it needs to come
quickly, what should we ethically do now to accelerate the process?

One possible short-term response to both questions: ACM can choose to
voluntarily impose a price on its emissions using carbon offsets.
\begin{itemize}
\item (What they are.)
\item giving ourselves a ``price signal''
is an effective way to begin raising awareness and changing organizational
behavior.
\item Lots of organizations (universities, companies) are doing this; some
ACM conferences are starting to experiment with voluntary carbon offset fees
at registration time. 
\item There are also some significant
concerns.
\item Concern:
Organizations are not set up to do it.  Our experience with getting
ACM to introduce an offsetting policy shows that it requires significant
effort.
\item Concern: It is hard to choose the price (right now there is an oversupply), and
setting it too low leads to complacency --- it doesn't really bite, so it
doesn't change people's behavior
\item This issue leads many people to argue that offsets "let people off the hook" and encourage
inaction, but we believe it is consistent with encouraging action -- again,
if offsets are priced right.
\end{itemize}

>>>>

\title{\Huge Conferences in the Era of Expensive Carbon}
\author{Benjamin C. Pierce (University of Pennsylvania) \\
Michael Hicks (University of Maryland) \\
Crista Lopes (University California, Irvine) \\
Jens Palsberg (University of California, Los Angeles)}

@@

A broad scientific consensus warns us that human
emissions of carbon dioxide and other greenhouse gases are warming the earth.
% (Figure \ref{fig:warming}).
This is not a problem to be left to future
generations: we are up against it now. The UN's Intergovernmental
Panel on Climate Change (IPCC) says a 40\% decrease in
emissions by 2030 is needed to avoid irreversible damage~\cite{IPCC-report}.
% (Figure~\ref{fig:IPCC-cover}).
Achieving reductions on this scale will
require urgent and sustained commitment at all levels of society, including
large organizations like universities, companies, and scientific societies
like ACM.

Indeed, scientific societies arguably have an especially important role to
play, since one of their main activities is organizing international
conferences, which produce significant greenhouse gas emissions,
particularly from air travel.   A single round-trip flight from
Philadelphia to Paris emits the equivalent of about 1.7 tons of carbon
dioxide (\COtwoE) per passenger~\cite{carboncalculator}\iflater\bcp{Trim
  citation?}\fi---a substantial fraction of the {\em total} yearly emissions
for an average resident of the US (16.5 tons) or Europe (7
tons)~\cite{emissions}.  \iflater\bcp{For comparison, ...  (put in some
  figures about the footprints of a few other things that people might think
  of reducing---driving, eating meat, upgrading their fridge...)}\fi{}
Moreover, these emissions have no near-term 
technological fix, since jet fuel is difficult to replace with
renewable energy sources~\cite{elec-air}. 
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

Conferences obviously also create significant value.  Science is a
fundamentally social process, and the conference system accelerates research
through high-bandwidth interaction, direct dissemination of results, network
building, and serendipitous cross-fertilization.

This leads to a puzzle: How do we weigh the environmental harm caused by
conference travel against the social benefits of the scientific progress
that comes from conferences?  It seems an extremely difficult question to
answer: the harm and the compensating benefits cannot be directly
compared---they are apples and oranges.  
%
Fortunately, there is a well-established mechanism for comparing apples and
oranges (indeed, for comparing almost anything): the magic of markets.  By
mapping choices and preferences in disparate domains to a common scale
(namely money), economic markets can be viewed as massively distributed
computations that balance costs an benefits across vastly different
activities.\bcp{Give a simple, concrete example.}  For this reason, many
policy experts advocate addressing climate change through some form of
carbon pricing, which places a monetary cost on greenhouse gas emissions,
thus giving a market signal that changes aggregate behavior.\bcp{See
  https://kleinmanenergy.upenn.edu/policy-digests/why-carbon-pricing-falls-short
  for lots of good phrasing and citations.}  Exactly how the price of carbon
should be set is a challenging policy question, but it has a clear goal: The
cost should be set so that the world arrives at zero emissions within the
timeframe that climate scientists tell us is safe.

In the long term, a carbon price should be uniform across society---i.e., it
should be imposed by governments.  Indeed, many countries and states (e.g.,
California) have already adopted carbon pricing policies; more will follow.
%
Moreover, although the current market price of carbon is still quite low
(around \$15 per ton), experts agree that this will need to rise
significantly over time to achieve the required effect.  Bottom line: We are
entering an era in which carbon is expensive.

Expecting that this era is coming, what should we rationally do now to
ensure that ACM, and computer-science research broadly, continue to thrive
in a world where air travel is more expensive?  Believing that it needs to
come as quickly as possible to avert disaster, what can we do to accelerate
the process?

A possible short-term response to both questions involves {\em carbon
  offsets}---a tool by which organizations (and individuals) can choose to
voluntarily impose a price on their own emissions.
%
% What they are:
In a nutshell, the idea would be that a conference could raise funds from
registration fees or corporate sponsors and give them, via a carbon offset
{\em vendor}, to someone who will use them to fund an activity (e.g.,
installing methane capture devices on landfill sites or buying efficient
cookstoves to replace open-fire cooking in developing areas) that will avoid
emitting an amount of greenhouse gases equivalent to the emissions from
participants' travel to the conference.
%
\begin{itemize}
\item giving ourselves a ``price signal''
is an effective way to begin raising awareness and changing organizational
behavior.
\item Lots of organizations (universities, companies) are doing this; some
ACM conferences are starting to experiment with voluntary carbon offset fees
at registration time. 
\item There are also some significant
concerns.
\item Concern:
Organizations are not set up to do it.  Our experience with getting
ACM to introduce an offsetting policy shows that it requires significant
effort.
\item Concern: It is hard to choose the price (right now there is an oversupply), and
setting it too low leads to complacency --- it doesn't really bite, so it
doesn't change people's behavior
\item This issue leads many people to argue that offsets "let people off the hook" and encourage
inaction, but we believe it is consistent with encouraging action -- again,
if offsets are priced right.
\end{itemize}
